Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-26 paper/paper.tex
pass on 5.1.2

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.}
  \centering
  \begin{tabular}{lllp{11cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Category} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth} \\
    \midrule     
    Ping Scan & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\

@@

Table~\ref{table:attacks} shows the attacks we considered in our
evaluation. These attacks cover different protocols and features of
\tname{}. Column ``Name'' shows the name of the attack, column
``Protocol'' shows the protocol name, and column ``Ground Truth''
shows the options of the rule that captures the attack.

\subsubsection{Malicious traffic.}To reproduce the malicious traffic
for http and tcp attacks, we configured an http server (Apache2) in
one machine and Nikto2~\cite{nikto} in a different machine. Nikto is
an open source tool with a catalog of scripts to reproduce
vulnerabilities. It is important to note that this setup is \emph{not}
necessary if \emph{pcap} files---pcap is the input format used by
\tname{}---for the attack already exist. As explained in
Section~\ref{sec:minimization}, pcap files abstract network
messages~\cite{pcap} and can be used to avoid network traffic during
rule synthesis. We proceeded as follows to obtain pcap files for the
the attacks we analyzed: we 1)~instantiated the script (\eg{}, defined
the IP of the web server), 2)~executed the Nikto2 script, 3)~and
monitored the traffic with the Wireshark network
monitor~\cite{wireshark-net-monitor}. The Wireshark tool records the
traffic and reports pcap files on output for replaying. The only
difference in the setup of non-http non-tcp attack is in the first
step of this procedure. For example, we used the hping3~\cite{hping3}
command to reproduce the ICMP SYN Flood attack: \CodeIn{hping3 -d 80
  -w 512 -S -p 80 --flood --rand-source <IP>}. This command alone
reproduces a flooding attack on a given machine
(Section~\ref{sec:dos}).\Mar{what about the variants? where do they
  come from? this is confusing as the narrative above suggests the
  attack is hardcoded in the tool.}\Luc{O trafego de um ataque 
  normalmente possui varios pacotes, entre esses varios pacotes alguns 
  deles possuem um padrao. As regras sao criadas descrevendo esse padrao,
  com isso um unico ataque pode disparar varios alertas.

  Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
  so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
  datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
  Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
  (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
  eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
  adicionar nas referencias: 
  https://www.netresec.com/?page=PcapFiles
}

@@

\Mar{Explique resumidamente o que cada ataque listado na tabela 2 faz...}

>>>>

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.\Mar{adicionar coluna ``Source''}}
  \centering
  \begin{tabular}{lllp{11cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Category} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & Active Recon. & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & DoS & TCP & \textbf{flags}: S,12; \textbf{threshold}: type both, track by\_dst, count 5000, seconds 5;\\

@@

\subsubsection{Attacks.}Table~\ref{table:attacks} shows the attacks we considered in our
evaluation. These attacks cover different protocols and features of
\tname{}. Column ``Name'' shows the name of the attack, column
``Protocol'' shows the protocol name, and column ``Ground Truth''
shows the options of the rule that captures the attack.\Mar{Explicar o
  rationale para esta selecao de ataques!}

%% It is important to note that this setup is \emph{not}
%% necessary if \emph{pcap} files for the attack already exist. We
%% proceeded as follows to obtain pcap files for the the attacks we
%% analyzed: we 1)~instantiated the script (\eg{}, defined the IP of the
%% web server), 2)~executed the script, 3)~and monitored the traffic with
%% the Wireshark network monitor. The
%% Wireshark tool records the traffic and reports pcap files on output
%% for replaying it.

\subsubsection{Malicious traffic.}~For some of the attacks listed on
Table~\ref{table:attacks}, we obtained pcap files directly from
trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). Recall that \tname\ takes pcap files on
input. As explained in Section~\ref{sec:minimization}, these files
abstract network messages~\cite{pcap} and can be used to avoid network
traffic during rule synthesis. In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool to reproduce
attacks. To reproduce the http and tcp attacks listed on
Table~\ref{table:attacks}, we configured an http server (Apache2) in
one machine and ran the security tool in a different
machine. Simultenously, we ran WireShark~\cite{wireshark-net-monitor}
to record the network traffic and Suricata to analyze the traffic. We
configured \suri\ with all rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and ran Nikto2 with default
parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved in each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that there are multiple instances of the same attack
encoded in Nikto2. RQ3 evaluates the importance of these variants to
\tname{}'s performance. Finally, Nikto2 only supports tcp and http
attacks.  We used the hping3~\cite{hping3} to reproduce the Ping Scan
icmp attack (see Section~\ref{sec:dos}).  More precisely, we used the
following hping3 command to reproduce that attack: \CodeIn{hping3 -d
  80 -w 512 -S -p 80 --flood --rand-source <IP>}.

%% \Mar{what about the variants? where do
%%   they come from? this is confusing as the narrative above suggests
%%   the attack is hardcoded in the tool.}\Luc{O trafego de um ataque
%%   normalmente possui varios pacotes, entre esses varios pacotes alguns
%%   deles possuem um padrao. As regras sao criadas descrevendo esse
%%   padrao, com isso um unico ataque pode disparar varios alertas.

%%   Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
%%   so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
%%   datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
%%   Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
%%   (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
%%   eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
%%   adicionar nas referencias: 
%%   https://www.netresec.com/?page=PcapFiles
%% }

@@
